---
title: "Biostimulatory analysis tool"
output: 
  html_document
self_contained: yes
runtime: shiny
---

<a href="https://github.com/SCCWRP/BioStimApp/" class="github-corner" aria-label="View source on Github"><svg width="80" height="80" viewBox="0 0 250 250" style="fill:#64CEAA; color:#fff; position: absolute; top: 0; border: 0; right: 0;" aria-hidden="true"><path d="M0,0 L115,115 L130,115 L142,142 L250,250 L250,0 Z"></path><path d="M128.3,109.0 C113.8,99.7 119.0,89.6 119.0,89.6 C122.0,82.7 120.5,78.6 120.5,78.6 C119.2,72.0 123.4,76.3 123.4,76.3 C127.3,80.9 125.5,87.3 125.5,87.3 C122.9,97.6 130.6,101.9 134.4,103.2" fill="currentColor" style="transform-origin: 130px 106px;" class="octo-arm"></path><path d="M115.0,115.0 C114.9,115.1 118.7,116.5 119.8,115.4 L133.7,101.6 C136.9,99.2 139.9,98.4 142.2,98.6 C133.8,88.0 127.5,74.4 143.8,58.0 C148.5,53.4 154.0,51.2 159.7,51.0 C160.3,49.4 163.2,43.6 171.4,40.1 C171.4,40.1 176.1,42.5 178.8,56.2 C183.1,58.6 187.2,61.8 190.9,65.4 C194.5,69.0 197.7,73.2 200.1,77.6 C213.8,80.2 216.3,84.9 216.3,84.9 C212.7,93.1 206.9,96.0 205.4,96.6 C205.1,102.4 203.0,107.8 198.3,112.5 C181.9,128.9 168.3,122.5 157.7,114.1 C157.9,116.9 156.7,120.9 152.7,124.9 L141.0,136.5 C139.8,137.7 141.6,141.9 141.8,141.8 Z" fill="currentColor" class="octo-body"></path></svg></a><style>.github-corner:hover .octo-arm{animation:octocat-wave 560ms ease-in-out}@keyframes octocat-wave{0%,100%{transform:rotate(0)}20%,60%{transform:rotate(-25deg)}40%,80%{transform:rotate(10deg)}}@media (max-width:500px){.github-corner:hover .octo-arm{animation:none}.github-corner .octo-arm{animation:octocat-wave 560ms ease-in-out}}</style>

```{r setup, warning = F, message = F, cache = T, echo = F}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.pat = 'fig/')

# packages
library(tidyverse)
library(shiny)
library(gridExtra)
library(shinyWidgets)
library(rsconnect)
library(DT)

# data
data(thrsdat)
data(bsbidat)
data(tllydat)

# unique bs variables
unibs <- tllydat %>% 
  pull(BSPretty) %>% 
  unique

# unique bi goals
unigl <- tllydat %>% 
  pull(BIgoal) %>% 
  unique

# unique bi response
unirs <- tllydat %>% 
  pull(Response) %>% 
  unique

# unique prob of meeting goal
unipr <- tllydat %>% 
  filter(!Prob %in% 'p50') %>% 
  pull(Prob) %>% 
  unique
```

For internal SCCWRP/Waterboard use only for review of draft biostimulatory analyses. These are preliminary results, and are subject to change. Do not distribute. For questions, contact [Raphael Mazor](mailto:raphaelm@sccwrp.org) or [Marcus Beck](mailto:marcusb@sccwrp.org).

Select the biostimulatory variables (one to all), biontegrity indices (one to all), one biointegrity goal, and one probability of meeting the goal.
```{r inputs}
column(12, 
       
       column(6, 
              pickerInput("selbs", label = 'Select biostim variable(s):', choices = unibs, multiple = TRUE,
                          selected = unibs
                                    )
       ),
       
       column(6, 
              pickerInput("selrs", label = 'Select biointegrity response(s):', choices = unirs, multiple = TRUE,
                          selected = unirs
                                    )
       )
)

column(12, 
       
       column(6, 
              selectInput("selgl", label = 'Select biointegrity goal:', choices = unigl, selected = 'Ref10'
                                    )
       ),
       
       column(6, 
              selectInput("selpr", label = 'Select probability of meeting goal:', choices = unipr, selected = 'p90'
                                    )
       )
       
)
```

```{r reactives}
# filter bi thresholds by selection
thrsrct <- reactive({
  
  # input
  selgl <- input$selgl
  selrs <- input$selrs
  
  out <- thrsdat %>% 
    filter(BIgoal %in% selgl) %>% 
    filter(Response %in% selrs) %>% 
    unique
  
  return(out)
  
})

# get all combinations from 0 to n for selected bs variables
cmbsrct <- reactive({

  # input
  trgtrct <- trgtrct()
  
  # validated bs variables
  valbs <- trgtrct$BSPretty
  
  # get all combos as nested list
  # unmet is '', met is 'X'
  out <- 0:length(valbs) %>% 
    sapply(function(m) combn(m, x = valbs, simplify = F), simplify = F) %>% 
    do.call('c', .) %>% 
    enframe %>% 
    unnest %>% 
    group_by(name) %>% 
    mutate(
      cnt = length(name),
      trg = ''
      ) %>% 
    ungroup %>% 
    spread(value, trg, fill = 'X')
  
  # all met add manually
  torbind <- tibble(
    name = 1, 
    cnt = 0, 
    bsvar = !! valbs,
    bsval = 'X'
    ) %>% 
    spread(bsvar, bsval)
  
  # join all
  out <- out %>% 
    rbind(torbind, .) %>% 
    mutate(
      cnt = max(cnt) - cnt # cnt is total passing as inverse
    ) %>% 
    mutate(cnt = as.integer(cnt)) 
  
  return(out)
    
})

# relevant targets for selection
trgtrct <- reactive({
  
  # input
  selbs <- input$selbs
  selrs <- input$selrs
  selgl <- input$selgl
  selpr <- input$selpr

  # filter by selection
  # order by minimum target estimate
  out <- tllydat %>% 
    filter(
      BSPretty %in% selbs &
      Response %in% selrs &
      BIgoal %in% selgl &
      Prob %in% selpr
    ) %>% 
    select(BSPretty, Response, Est) %>% 
    group_by(BSPretty) %>% 
    arrange(BSPretty, Est) %>% 
    mutate(
      rnk = 1,
      rnk = cumsum(rnk),
      rnk = as.integer(rnk)
    ) %>% 
    filter(rnk == 1) %>% 
    ungroup %>% 
    select(BSPretty, Est)
   
  return(out)
  
})

# filter observed site data by selection, get bi category and bs pass/fail
bsbirct <- reactive({

  # input
  selrs <- input$selrs
  selgl <- input$selgl
  trgtrct <- trgtrct()
  thrsrct <- thrsrct()
  cmbsrct <- cmbsrct()
  
  req(nrow(trgtrct) > 0)
  
  # validated bs variables
  valbs <- trgtrct$BSPretty
  
  # filter observed, join with threshold and bs target data, get pass fail
  alltlly <- bsbidat %>% 
    filter(
      BSPretty %in% valbs &
      Response %in% selrs
      ) %>% 
    left_join(trgtrct, by = 'BSPretty') %>% 
    left_join(thrsrct, by = 'Response') %>% 
    mutate(
      bsfail = ifelse(bsval > Est, '', 'X'),
      bipass = bival >= glval
    ) %>% 
    group_by(MasterID, SampleDate) %>% 
    mutate(
      bipassall = ifelse(any(!bipass), FALSE, TRUE)
    ) %>% 
    select(MasterID, SampleDate, bipassall, DevSet, BSPretty, bsfail) %>% 
    unique
  
  # get pass fail counts, total passing in each group summarized
  sumtlly <- alltlly %>% 
    group_by(MasterID, SampleDate, DevSet) %>% 
    filter(!anyNA(bipassall)) %>% # remove sites with incomplete bi data
    filter(!anyNA(bsfail)) %>% # remove sites with incomplete bs data
    spread(BSPretty, bsfail) %>% 
    group_by_at(vars(one_of(c('DevSet', valbs)))) %>% 
    summarise(
      bipasstot = sum(bipassall),
      bipassall = length(bipassall),
      bipassprp = bipasstot / bipassall,
      bipassprp = round(bipassprp, 2)
    ) %>% 
    select(-bipasstot) %>% 
    gather('sumvar', 'sumval', bipassall, bipassprp) %>% 
    ungroup %>% 
    unite('sumvar', DevSet, sumvar, sep = '_') %>% 
    spread(sumvar, sumval, fill = 0)
  
  # join with all possible combos from cmbsrct
  # give it pretty names
  out <- sumtlly %>% 
    left_join(cmbsrct, ., by = valbs) %>% 
    replace(., is.na(.), 0) %>% 
    mutate(
      Cal_bipassall = as.integer(Cal_bipassall),
      Val_bipassall = as.integer(Val_bipassall)
    ) 

  return(out)
      
})

```

This table shows the lowest validated biostimulatory targets for the selections above. Some biostimulatory variables may not be shown if the targets were not valided.   
```{r output}
renderTable({
  
  # input
  trgtrct <- trgtrct()
  
  # labels with units
  labs <- tibble(
    BSPretty = c('Total N', 'Total P', 'Chl-a', 'AFDM', '% cover'),
    BSlabs = c('Total N (mg/L)', 'Total P (mg/L)', 'Chla-a (mg/m2)', 'AFDM (mg/cm2)', '% macroalgae cover')
  )
  
  # slight formatting 
  trgtrct <- trgtrct %>% 
    left_join(labs, by = 'BSPretty') %>% 
    select(BSlabs, Est) %>% 
    rename(
      `Biostimulatory variable` = BSlabs,
      `Lowest valided target` = Est
    )
    
  if(nrow(trgtrct) == 0)
    trgtrct <- 'No targets were validated'
  
  trgtrct
  
  })
```

This table shows the site summaries for all combinations of meeting or not meeting the biostimulatory targets.  Targets tha are met are designated with "X".  The totals indicate the number of sites in each category and the proportion of those sites for which all selected biointegrity goals are met.  Totals are separated by calibration and validation sites.  Rows with ten or more calibration sites are in bold. Rows can be sorted using the selection arrows next to each column name (e.g., sort by a biostimulatory variable to see other variables that were not met).
```{r}
output$sumtab <- renderDataTable({
  
  # input
  bsbirct <- bsbirct()
  
  # rows where number of cal sites > 10
  tobld <- bsbirct %>% 
    pull(Cal_bipassall) %>% 
    ">=" (10) %>% 
    which
  
  # table data
  totab <- bsbirct %>% 
    select(-name)%>% 
    rename(
      `Total biostim targets met` = cnt, 
      `Total sites (cal)` = Cal_bipassall,
      `Proportion meeting selected biointegrity targets (cal)` = Cal_bipassprp,
      `Total sites (val)` = Val_bipassall,
      `Proportion meeting selected biointegrity targets (val)` = Val_bipassprp
    )

  # tab output
  datatable(totab, 
            options = list(dom = 't', pageLength = 32, fixedHeader = TRUE),
            extensions = "FixedHeader"
            ) %>% 
    formatStyle(columns = c(2:6), 'text-align' = 'center') %>%
    formatStyle(0,
      target = "row",
      fontWeight = styleEqual(tobld, rep("bold", length(tobld)))
    )
  
})
 
dataTableOutput('sumtab')
```

